---
title: "fitbit0analysis"
author: "Deepak Grewal"
date: "4/1/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 
```{r, message=FALSE}
library(dplyr)
library(tidyverse)
 

```

## Reading in data
```{r, message=FALSE}
 

activity <- read_csv("activity.csv")

```

## analyse Activity
What is mean total number of steps taken per day?
```{r, message=FALSE}
 
# derive mean and median steps per day 
summSteps <- activity %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE)) %>% ungroup() %>% 
  mutate(allSteps = sum(TotalSteps),meanPerDay= mean(TotalSteps), medianPerDay= median(TotalSteps))

summSteps$meanPerDay[1]

```
 
 

Calculate the total number of steps taken per day

```{r, message=FALSE}

# calculate total steps on each day
TotalSteps <- activity %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE))
```

 Make a histogram of the total number of steps taken each day
 Use the data containing total steps
 
```{r}

# use histogram to display steps
TotalSteps %>%  ggplot(aes(x= TotalSteps)) + 
  geom_histogram(bins = 60, fill='lightgreen')+
  labs(x= "Total Steps On a Day" , y= "Number of Days")+
  theme_classic()
```
 

Calculate and report the mean and median of the total number of steps taken per day

```{r , message=FALSE}

# derive mean and median steps per day 
summSteps <- activity %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE)) %>% ungroup() %>% 
  mutate(allSteps = sum(TotalSteps),meanPerDay= mean(TotalSteps), medianPerDay= median(TotalSteps))



```

Mean steps Per Day
```{r }
summSteps$meanPerDay[1]
```
Median steps Per Day
```{r }
summSteps$medianPerDay[1]
```

What is the average daily activity pattern?
Make a time series plot  for the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

first get average steps for the intervals
```{r ,message=FALSE}
# derive mean and median steps per 5 minute interval 
summStepsInt <- activity %>%  group_by(interval) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE),meanPerInt= mean(steps, na.rm = TRUE))
```

Plot the data
```{r}
#plot average steps across the time points
summStepsInt %>%  ggplot(aes(x= interval, y= meanPerInt)) + geom_line(color='blue') +
  theme_classic()
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Maximum Steps Interval:
```{r}
# get the maximum average step time interval

maxPerInt <- summStepsInt %>% filter(meanPerInt== max(meanPerInt))

maxPerInt$interval
```


## Impute Missing Data and Analyse activity

Following Number of records are missing

```{r}
#get total records missing records
missing_rec <- activity %>% filter(is.na(steps) | is.na(date) | is.na(interval) ) %>% count()
missing_rec$n
```

Replace missing with median values 
```{r, message=FALSE}
# replace the missing steps for the interval by median steps for that interval
#1- calculate the median steps per interval
summStepsInt2 <- activity %>%  group_by(interval) %>% 
  summarise(medianPerInt= median(steps, na.rm = TRUE)) 
#2- replace missing with median value

activityImputed <- activity %>% left_join(summStepsInt2 , by= c("interval")) %>% 
  mutate(steps=if_else(is.na(steps),medianPerInt, steps )) %>% select(-c(medianPerInt))

```
histogram of Imputed steps
```{r message=FALSE}

# calculate total steps on each day
TotalSteps <- activityImputed %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE)) 

# use histogram to display steps
TotalSteps %>%  ggplot(aes(x= TotalSteps)) + 
  geom_histogram(bins = 60, fill='lightgreen')+
  labs(x= "Total Steps On a Day" , y= "Number of Days")+
  theme_classic()
```
Mean and Median of imputed data

```{r, message=FALSE}
# derive mean and median steps per day 
summSteps <- activityImputed %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE)) %>% ungroup() %>% 
  mutate(allSteps = sum(TotalSteps),meanPerDay= mean(TotalSteps), medianPerDay= median(TotalSteps))

summSteps$meanPerDay[1]
summSteps$medianPerDay[1]

```
 As you can see above median Did not change(as median was used to replace imputed) , but mean changed
 
## Analyse Weekend Vs Weekdays

1. add factor column for type of Day- Weekday or Weekend and derive data for plot
```{r, message=FALSE}

# add factor
activityDay <- activity %>% mutate(day=as.factor(if_else(
  weekdays(date) %in% c("Monday" ,   "Tuesday"  , "Wednesday", "Thursday" , "Friday"), "Weekday", "Weekend" )))

 
# derive mean and median steps per 5 minute interval 
summStepsInt <- activityDay %>%  group_by(interval, day) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE),meanPerInt= mean(steps, na.rm = TRUE)) 
```

2. Display plot

```{r}

#plot average steps across the time points
summStepsInt %>%  ggplot(aes(x= interval, y= meanPerInt, color=day)) + geom_line() +
  theme_classic()

```


