---
title: "fitbit0analysis"
author: "Deepak Grewal"
date: "4/1/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 Reading in data
```{r}
library(dplyr)
library(tidyverse)

activity <- read_csv("activity.csv")

```
What is mean total number of steps taken per day?
```{r}
 
# derive mean and median steps per day 
summSteps <- activity %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE)) %>% ungroup() %>% 
  mutate(allSteps = sum(TotalSteps),meanPerDay= mean(TotalSteps), medianPerDay= median(TotalSteps))

summSteps$meanPerDay[1]

```
 
 

Calculate the total number of steps taken per day

```{r}

# calculate total steps on each day
TotalSteps <- activity %>%  group_by(date) %>% 
  summarise(TotalSteps= sum(steps, na.rm = TRUE))
```

 Make a histogram of the total number of steps taken each day
 Use the data containing total steps
 
```{r}

# use histogram to display steps
TotalSteps %>%  ggplot(aes(x= TotalSteps)) + 
  geom_histogram(bins = 60, fill='lightgreen')+
  labs(x= "Total Steps On a Day" , y= "Number of Days")+
  theme_classic()
```
 

Calculate and report the mean and median of the total number of steps taken per day

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
